### 2021-10-28 paper/SupMat.Rnw
finished revising notation in SupMat

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}
\usepackage{mathrsfs}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{accents}

%new commands and so on
\DeclareRobustCommand{\bbone}{\text{\usefont{U}{bbold}{m}{n}1}}

\DeclareMathOperator{\EX}{E}% expected value
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\VarX}{var}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\CovX}{cov}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\Prob}{P}

\newcommand{\sC}{\mathscr{C}} %for a script C like Ellner et al (2016), Ecol Lett
\newcommand{\bs}{\backslash}

%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decide not to use)
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\olep}{\overline{\epsilon}}

%external documents
\externaldocument[MT-]{Paper}

%header material for paper
\title{Relationships in the extremes and their influence on competition and coexistence: Supporting information}
\date{}

\author[a]{Pimsupa Jasmin Albert}
\author[a,b]{Daniel C. Reuman}

\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}

%***Need to indicate corresponding authors

\begin{document}
\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
#system(paste("biber", sub("\\.Rnw$", "", current_input())))

#place R code here for loading in necessary variables from Results
#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
#allregres<-matrix(rnorm(10),2,5)
M <- readRDS(file="../results_numeric/M.RDS")
rho <- readRDS(file="../results_numeric/rho.RDS")
se3qij <- readRDS(file="../results_numeric/fig3qijmaxse.RDS")
@

\maketitle

\tableofcontents
\listoffigures

%probably want table of contents, figure and table lists

\section{Scaling factors}\label{SIsect:scaling_factors}

Following \cite{Ellner_2016}, we define
\begin{equation}
\sC_j = -r_j(\overline{E}_j,C_j),
\end{equation}
and 
\begin{equation}
\sC_{j \bs i} = -r_j(\overline{E}_j,C_{j \bs i}),
\end{equation}
i.e., $\sC_{j \bs i}$ is the negative of the growth rate of species $j$ when $E_j$ is at its mean value
and when $i$ is at negligible abundance. Then $q_{ij}$ is defined by computing
\begin{equation}
\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}},
\end{equation}
and evaluating it at $\sC_{j \bs i}=0$ [see \cite{Chesson_1994} and \cite{Ellner_2016} for background
on this definition]. To make sense of this definition, it must be possible to write $\sC_{i \bs i}$, explicitly 
or implicitly, uniquely as a function of $\sC_{j \bs i}$ \citep{Ellner_2016}.

For the lottery model, for $i \neq j$, we have
\begin{align}
\sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
\sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
\sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
\end{align}
Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
$B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it calculate $q_{ij}$,
but we instead compute $q_{ij}$ through implicit differentiation:
\begin{align}
\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
&= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
\end{align}
Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
\begin{align}
q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
&= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
\end{align}
as stated in the main text.

%***DAN: Change the below paragraph if needed. Some reasons it might be needed: Steve joins as a coauthor;
%Steve reads this and thinks it makes sense to change it because something is factually or presentationally 
%sub-optimal; or Dan or Jasmin actually develops a more rigorous understanding of scaling factors and based
%on that thinks something here is sub-optimal. Anyway, the "do both" approach seems like it can't go wrong,
%especially if Steve's prediction is correct that it won't make much difference.
%***DAN: You probably need to spend a bit more time with the cited papers to better understand this. 
We also use $q_{ij}=1$ as an alternative value, based on communication with Steve Ellner, for the following reasons.
First, as discussed in the supporting information of \cite{Ellner_2016}, and as further elaborated by 
\cite{Barabas_2018}, it takes very special circumstances for the scaling factors to exist and be unique, 
especially with size-structured models. Second, in empirically parameterized models, scaling factor values are 
often not robust to arbitrary modelling choices. 
Although the lottery model is not size structured and we do not empirically parameterize it, the above shortcomings of 
scaling factors, generally, suggest we should find alternatives to their use. The alternative 
suggested by Ellner for empirical applications is to do a series of pairwise invader-resident comparisons, each with an 
equal weighting of invader and resident $r$ values; for our context this amounts to setting scaling factors to $1$. 
We consider both approaches because the above-mentioned shortcomings do not apply to our case, but it seems 
possible that future applications will often set scaling factors to $1$, and we want our ideas to also be considered 
by researchers making that choice.

\section{Mechanisms other than storage effects}\label{SIsect:other_mechanisms}

We first review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
original decompositions of \cite{Chesson_1994}; and then we include our extensions of the decomposition.

We begin by decomposing 
the invasion growth rate,
\begin{equation}
\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})].
\end{equation}
The invasion growth rate with both $E_i$ and $C_{i \bs i}$ constant at their means is 
\begin{align}
\epsilon_i^0 = r_i(\overline{E}_i,\overline{C}_{i \bs i}).
\end{align}
The additional contribution of variation in $E_i$ to the invasion growth rate is then
\begin{equation}
\olep_i^E = \E[r_i(E_i,\overline{C}_{i \bs i})]-\epsilon_i^0,
\end{equation}
whereas the additional contribution of variation in $C_{i \bs i}$ is 
\begin{equation}
\olep_i^C=\E[r_i(\overline{E}_i,C_{i \bs i})]-\epsilon_i^0.
\end{equation}
We then define
\begin{align}
\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0.
\end{align}
Thus we have the primary decomposition 
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC}.
\end{equation}
But some of $\olep_i^{EC}$ comes from the fact that $E_i$ and $C_{i \bs i}$ are
both varying, and some comes from covariation in these two quantities. So we next separate these
contributions in a secondary decomposition of $\olep_i^{EC}$.

Let
\begin{equation}
\olep_i^{(E\shp C)}=\E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0
\end{equation}
and let
\begin{equation}
\olep_i^{(EC)} = \olep_i^{EC}-\olep_i^{(E\shp C)},
\end{equation}
so
\begin{equation}
\olep_i^{(EC)}=\E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)].
\end{equation}
The parentheses in the superscripts of $\olep_i^{(EC)}$ and $\olep_i^{(E\shp C)}$
indicate that we are talking about a term in the secondary decomposition, i.e., in the decomposition of 
$\olep_i^{EC}$. We now have
\begin{align}
\olr_{i \bs i} &= \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC} \\
\olep_i^{EC} &= \olep_i^{(E\shp C)}+\olep_i^{(EC)},
\end{align}
i.e.,
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C +\olep_i^{(E\shp C)}+\olep_i^{(EC)}. 
\end{equation}
The term $\olep_i^{(EC)}$ is due to covariation between $E_i$ and $C_{i \bs i}$, whereas the term 
$\olep_i^{(E\shp C)}$ is due to variation \emph{per se} in both $E_i$ and $C_{i \bs i}$, after the effects of 
covariation itself are removed. 
The term $\olep_i^{(EC)}$ relates to storage effects, as we will see.

As in the main text, we can now define the tertiary decomposition
\begin{align}
\olep_i^{(EC)} &= \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \\
&= \left[ \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)] \right]+\left[ \E[r_i(E_i^\nat,C_{i \bs i}^\nat)]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \right] \\
&= \olep_i^{[EC]}+\olep_i^{[E \nat C]}.
\end{align}
The brackets in the superscripts of $\olep_i^{[EC]}$ and $\olep_i^{[E \nat C]}$ indicate that we are talking about a term in the
tertiary decomposition, i.e., in the decomposition of $\olep_i^{(EC)}$. 
The term $\olep_i^{[EC]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on the invasion growth rate,
and the the term $\olep_i^{[E \nat C]}$ corresponds to $EC$ correlations \emph{per se}, after the removal of asymmetries of tail association.
The symbol $\nat$ denotes the removal of asymmetries of tail association. We therefore have 
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[E \nat C]}+\olep_i^{[EC]},\label{eq:decomp_overall_IGR}
\end{equation}
our overall decompsition of $r_{i \bs i}$.
The same derivation also applies to provide 
\begin{equation}
\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[E \nat C]}+\olep_j^{[EC]},\label{eq:decomp_overall_rjbsi}
\end{equation}
an analogous decomposition of $\olr_{j \bs i}$.

Combining the decompositions of $\olr_{i \bs i}$ and $\olr_{j \bs i}$ and assuming without loss of generality that $\mu_i \leq \mu_j$ so that
species $i$ is the weaker competitor and $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ can be regarded as a coexistence metric, we have
\begin{equation}
\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]}+\Delta_i^{[EC]},\label{eq:decomp_overall_delta}
\end{equation}
where $\Delta_i^0=\epsilon_i^0- q_{ij} \epsilon_j^0$;
$\Delta_i^E=\olep_i^E -q_{ij}\olep_j^E$ is the contribution to coexistence of environmental variation;
$\Delta_i^C=\olep_i^C -q_{ij}\olep_j^C$ is the contribution to coexistence of variation in competitive pressure;
$\Delta_i^{(E\shp C)}=\olep_i^{(E\shp C)} - q_{ij}\olep_j^{(E\shp C)}$ is the contribution to coexistence of 
variation \emph{per se} in both environment and competition, not including the effects of covariation of these quantities;
 $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$is the contribution to coexistence of $EC$ covariation \emph{per se},
not including the effects of asymmetric tail associations between environment and competition;
and $\Delta_i^{[EC]}=\olep_i^{[EC]} - q_{ij} \olep_j^{[EC]}$ is the contribution to coexistence of 
asymmetric tail associations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.

\section{Noise} \label{sect:noise}

We created two cases for $(b_1,b_2)$, left- and right-tail association. 
To generate $M$ points $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N\left( \vec{0}, \Sigma \right)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$. 
Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -|a_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$. 
We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. 
Perfect association in the left tails of the resulting distribution $\left(b_1,b_2\right)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value. 
For speed of computation, when samples $(b_1^{(i)},b_2^{(i)})$, $i=1,...,M$ were needed for distinct sets of values of the parameters $\mu_1$, $\mu_2$, and $\sigma$, we used the same $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values. Right-tail associated noise was generated in an analogous way.

When needed, noise with symmetric tail association was generated by taking a very large sample (\Sexpr{M}) 
of left-tail associated noise generated as 
described above, and then calculating the Pearson correaltion coefficient, $\rho$, of the sample.
Symmetric noise $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$ for $i=1,...,M$ was then generated 
from the bivariate normal distribution $N\left(\vec{0}, \Sigma \right)$, where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, 
and then we set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.

Overall, our version of the lottery model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether 
noise was generated from the symmetric, left- or right-tail associated cases.

\section{Mathematical simplifications for plotting}\label{sect:math_simplifications}

We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the lottery model. 
One can straightforwardly show from the formulas of SI section \ref{SIsect:other_mechanisms} that
\begin{equation}
E_i=B_i=\exp(b_i)
\end{equation}
\begin{equation}
C_i=\frac{B_1 N_1 + B_2 N_2}{\delta N}
\end{equation}
\begin{equation}
C_{i \bs i} = B_j/\delta=\exp(b_j)/\delta
\end{equation}
\begin{equation}
r_i=\ln(1-\delta+E_i/C_i)
\end{equation}
%IGR
\begin{align}
\olr_{i \bs i} &= \E[\ln(1-\delta+E_i/C_{i \bs i})] \\
&= \E[\ln(1-\delta+\delta B_i/B_j)] \\
&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]\label{eq:base_IGR}
\end{align}
\begin{equation}
\overline{E}_i = \exp(\mu_i + \sigma^2/2)
\end{equation}
\begin{equation}
\overline{C}_{i \bs i}=\exp(\mu_j+\sigma^2/2)/\delta
\end{equation}
%baseline
\begin{equation}
\epsilon_i^0 = \ln(1-\delta+\delta \exp(\mu_i - \mu_j))\label{eq:decomp:baseline}
\end{equation}
%contrib of variation in E
and
\begin{align}
\olep_i^E &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i+\mu_i)/\exp(\mu_j+\sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}
\end{align}
where $\tilde{b}_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%contrib of variation in C
\begin{align}
\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(b_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma \tilde{b}_j +\mu_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(-\sigma \tilde{b}_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 
\end{align}
where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
\begin{align}
\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(b_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i +\mu_i)/\exp(\sigma \tilde{b}_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}
\end{align}
where $(\tilde{b}_i, \tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$. Continuing,
%contrib of variation in E and C, per se  
\begin{align}
\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(b_i^\shp)/\exp(b_j^\shp))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i^\shp +\mu_i)/\exp(\sigma \tilde{b}_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}
\end{align}
where $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
%storage effects
\begin{equation}
\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}
\end{equation}
where $(\tilde{b}_i,\tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$, and 
$\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normal and independent. Continuing,
%contrib of ATA
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}
\end{equation}
where $(\ddot{b}_i,\ddot{b}_j)$ have the same (normal) marginals and covariance as $(b_i,b_j)$ but are 
bivariate normal. Thus
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
\end{equation}
where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
from expression (\ref{eq:decomp:storage_effects}), i.e., 
\begin{equation}
\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
\end{equation}
This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.

We next decompose $\olr_{j \bs i}$ for the lottery model. We have 
\begin{equation}
E_j = B_j = \exp(b_j)
\end{equation}
\begin{equation}
C_j=\frac{B_1 N_1 + B_2 N_2}{\delta N}
\end{equation}
\begin{equation}
C_{j \bs i} = B_j/\delta = \exp(b_j)/\delta
\end{equation}
\begin{equation}
r_j=\ln(1-\delta+E_j/C_j)
\end{equation}
\begin{align}
\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
&= 0
\end{align}
\begin{equation}
\overline{E}_j = \exp(\mu_j+\sigma^2/2)
\end{equation}
\begin{equation}
\overline{C}_{j \bs i} = \exp(\mu_j + \sigma^2/2)/\delta
\end{equation}
%baseline
\begin{align}
\epsilon_j^0 &= r_j(\overline{E}_j,\overline{C}_{j \bs i}) \\
&= \ln \left( 1-\delta+\delta \frac{\exp(\mu_j + \sigma^2/2)}{\exp(\mu_j + \sigma^2/2)} \right) \\
&= 0 \label{eq:decomp:baseline_j}
\end{align}
and
%contrib of variation in E
\begin{align}
\olep_j^E &= \E[r_j(E_j,\overline{C}_{j \bs i})]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta B_j/\overline{B}_j)]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma \tilde{b}_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma \tilde{b}_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}
\end{align}
where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%contrib of variation in C
\begin{align}
\olep_j^C &= \E[r_j(\overline{E}_j,C_{j \bs i})]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\overline{B}_j}{B_j} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(b_j)} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma \tilde{b}_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(-\sigma \tilde{b}_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}
\end{align}
where again $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
\begin{align}
\olep_j^{EC} &= \E[r_j(E_j,C_{j \bs i})]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta B_j/B_j)]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= -\olep_j^E-\olep_j^C-\epsilon_j^0.\label{eq:decomp:j:EC}
\end{align}
Continuing,
%contrib of variation in E and C, per se  
\begin{align}
\olep_j^{(E\shp C)} &= \E[r_j(E_j^\shp , C_{j \bs i}^\shp )]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{B_j^\shp}{B_j} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j^\shp)}{\exp(b_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma \tilde{b}_j^\shp + \mu_j)}{\exp(\sigma \tilde{b}_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}
\end{align}
where $(E_j^\shp , C_{j \bs i}^\shp)$ has the same marginals as $(E_j,C_{j \bs i})$ but has independent components;
$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; $b_j^\shp$ is distributed in the same
way as $b_j$ but is independent of it; and $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and 
independent of each other. Continuing,
%storage effects
\begin{align}
\olep_j^{(EC)} &= \E[r_j(E_j,C_{j \bs i})]-\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
&= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:pECp}
\end{align}
these results obtained from (\ref{eq:decomp:j:EC}) and (\ref{eq:decomp:j:EshpC}). Again, $\tilde{b}_j$ and 
$\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
\begin{align}
\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
&= 0 \label{eq:decomp:j:EnatC}
\end{align}
because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that
$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
\begin{align}
\olep_j^{[E \nat C]} &= \olep_j^{(EC)} - \olep_j^{[EC]} \\
&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}
\end{align}
where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.

This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.

Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
$\delta$, $\sigma$, and $\mu_i - \mu_j$, but not separately on $\mu_i$ or $\mu_j$. The terms of the decomposition
of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\mu_i$ or $\mu_j$ at all.
As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
\begin{align}
q_{ij} &= \frac{\exp(\mu_i)}{(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)} \\
&= \frac{\exp(-\mu_i)\exp(\mu_i)}{\exp(-\mu_i)[(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)]} \\
&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.\label{eq:decomp:qnot1}
\end{align}
These also either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
$\mu_i-\mu_j$. Thus each of the terms of the decomposition (\ref{MT-eq:full_decomp}) from the main text
depends only on $\mu_i-\mu_j$ and not separately on $\mu_i$ or $\mu_j$, as stated in the Methods section of the main text.

We now want to show that, for the lottery model, 
all elements of our decompositions (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), as well
as $q_{ij}$, and therefore all elements of our decomposition (\ref{eq:decomp_overall_delta}), are the same 
for our left- and right-tail associated cases, as was stated in Methods in the main text. Using the notation 
$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(b_i^{(l)}),\exp(b_j^{(l)})/\delta)$ and 
$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(b_i^{(r)}),\exp(b_j^{(r)})/\delta)$
for the left- and right-tail associated cases, respectively, we note that the marginal distributions of
$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and likewise the marginals of 
$(B_i^{(l)},B_j^{(l)}/\delta)$ are the same as those of $(B_i^{(r)},B_j^{(r)}/\delta)$.
So any component of (\ref{eq:decomp_overall_IGR}), or $q_{ij}$, will not depend on the choice
of our left- or right-tail associated noise cases as long as it depends only on marginal distributions.
But it is easy to see that $q_{ij}$ depends only on marginals, since $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$  
depends only on $\mu_i$, $\mu_j$ and $\delta$, and these are parameters of the marginal distributions of 
$(b_i,b_j)$. The alternative value $q_{ij}=1$ is constant, so also takes the same value for both our
left- and right-tail association noise cases. It is also easy to see from the definitions of $\epsilon_i^0$, 
$\olep_i^E$, $\olep_i^C$, and $\olep_i^{(E \shp C)}$ that these quantities only depend on marginals.

Whereas $\olep_i^{[E \nat C]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
more than just marginals, this quantity is still independent of the choice between our left- and right-tail-association 
cases because the bivariate random variables $(E_i^\nat , C_{i \bs i}^\nat )$ and 
$(E_i^\shp ,C_{i \bs i}^\shp)$ have had their tail association removed, and the only differences 
between $(E_i^{(l)},C_{i \bs i}^{(l)})$ and $(E_i^{(r)},C_{i \bs i}^{(r)})$ are in their tail association
properties (i.e., these two bivariate distributions have the same marginals and covariances).
In other words, $(E_i^\nat , C_{i \bs i}^\nat )$ is distributed in the same way for the left- and right-tail-association cases,
as is $(E_i^\shp ,C_{i \bs i}^\shp)$.

We now examine $\olep_i^{[EC]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

\section{The $\delta=1$ case}\label{sect:delta1}

%The i terms
Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
\begin{equation}
\epsilon_i^0 = \mu_i - \mu_j.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
\begin{align}
\olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
&= -\sigma^2/2.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
\begin{align}
\olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
&= \sigma^2/2.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
\begin{align}
\olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
&= 0.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
\begin{align}
\olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
&= 0.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
\begin{align}
\olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
&= 0.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
\begin{equation}
\olep_i^{[E \nat C]}=0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
\begin{equation}
\olep_i^{[EC]}=0.
\end{equation}
%The j terms
Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
\begin{equation}
\epsilon_j^0 = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
\begin{equation}
\olep_j^E = -\sigma^2/2.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
\begin{equation}
\olep_j^C = \sigma^2/2.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
\begin{equation}
\olep_j^{EC} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
\begin{equation}
\olep_j^{(E \shp C)} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
\begin{equation}
\olep_j^{(EC)} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
\begin{equation}
\olep_j^{[E \nat C]} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
\begin{equation}
\olep_j^{[EC]}=0.
\end{equation}

For $\delta=1$, the second value of $q_{ij}$ we considered, namely
$q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
equations using $q_{ij}=1$ gives
\begin{align}
\Delta_i^0 &= \mu_i - \mu_j \\
\Delta_i^E &= 0  \\
\Delta_i^C &= 0  \\
\Delta_i^{(E \shp C)} &=  0 \\
\Delta_i^{[E \nat C]} &= 0  \\
\Delta_i^{[EC]} &=  0. 
\end{align}
Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
without loss of generality).
So there is no robust coexistence in the $\delta=1$ case for our simple model.

\section{Efficient computation}\label{sect:efficient_computation}

We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the lottery model.

First, for a large integer $M$ (we used $M=\Sexpr{M}$),
carry out the following steps once.
\begin{enumerate}
\item Generate $M$ independent points $u^{(k)}$, $k=1,\ldots,M$, from a standard normal distribution.
\item Generate $M$ independent points $(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$, as in SI section \ref{sect:noise}, which are left-tail associated
and which have standard-normal marginals. 
\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(\tilde{\ddot{b}}_i^{(k)},\tilde{\ddot{b}}_j^{(k)})$, 
$k=1,\ldots,M$, from a bivariate normal distribution with component variances $1$ and Pearson correlation the same as 
$(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$.
\end{enumerate}
These data are stored and used repeatedly below. This provides computational efficiency because 
random number generation is computationally expensive.

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well

Next, proceed as follows to get estimates of the terms in the decomposition (\ref{eq:decomp_overall_rjbsi}). 
\begin{enumerate}
\item Use $\epsilon_j^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
\item Estimate $\olep_j^E$, following (\ref{eq:decomp:contrib_of_var_E_j}), as
$\widehat{\olep_j^E}=\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]-\epsilon_j^0$, an estimate
with standard error $\se \left( \widehat{\olep_j^E} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^C$, following (\ref{eq:decomp:contrib_of_var_C_j}), as
$\widehat{\olep_j^C}=\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]-\epsilon_j^0$, which has 
standard error $\se \left( \widehat{\olep_j^C} \right) = \sd_k [\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_j$ and $\tilde{b}_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
independent, so $\tilde{b}_j^\shp - \tilde{b}_j$ is normally distributed with mean $0$ and variance $2$. So we can 
estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:decomp:j:EshpC}), as 
$\widehat{\olep_j^{(E \shp C)}} = \mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0$.
The standard error of this expression is $\se \left(  \widehat{\olep_j^{(E \shp C)}} \right) = 
\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))-\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^{[E \nat C]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[E \nat C]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[E \nat C]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
\item Use $\olep_j^{[EC]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
\item Finally, use $\olr_{j \bs i} = 0$, again exact. 
\end{enumerate}
As for (\ref{eq:decomp_overall_IGR}), it is again easy to see that (\ref{eq:decomp_overall_rjbsi}) applies for the 
estimates as well as for the quantities being estimated.

Next, $q_{ij}$ can be taken (see SI section \ref{SIsect:scaling_factors}) to be, alternatively, $1$ or 
the quantity displayed in (\ref{eq:decomp:qnot1}). Hence the terms of the decomposition 
(\ref{eq:decomp_overall_delta}) can straightforwardly be etimated using our estimates elaborated 
above, as follows. There are two alternative versions of the estiamtors for this decomposition, corresponding 
to the alternative values for $q_{ij}$.
\begin{enumerate}
\item Compute $\Delta_i^0 = \epsilon_i^0 - q_{ij} \epsilon_j^0$, which is computed exactly, since 
the quantities described above for the components $\epsilon_i^0$, $\epsilon_j^0$ and $q_{ij}$ are exact, as opposed to being
estimates. Simplifying, since $\epsilon_j^0=0$, we have $\Delta_i^0 = \epsilon_i^0$.
\item Estimate $\Delta_i^E$ as $\widehat{\Delta_i^E}=\widehat{\olep_i^E}-q_{ij}\widehat{\olep_j^E}=
\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0-
q_{ij} \mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))],$ which has standard error 
$\se\left( \widehat{\Delta_i^E} \right)=\sd_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
q_{ij} \ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}.$
\item Estimate $\Delta_i^C$ as $\widehat{\Delta_i^C}=\widehat{\olep_i^C}-q_{ij}\widehat{\olep_j^C}=
\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-
\epsilon_i^0-q_{ij}\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))],$ which has standard error
$\se \left( \widehat{\Delta_i^C} \right)=\sd_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))-
q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\Delta_i^{(E \shp C)}$ as $\widehat{\Delta_i^{(E \shp C)}} = \widehat{\olep_i^{(E \shp C)}}-q_{ij} \widehat{\olep_j^{(E \shp C)}}=
\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0-
q_{ij}[\mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0].$
This simplifies to $\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-
q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\Delta_i^E}-\widehat{\Delta_i^C}-\Delta_i^0$, 
which can be straightforwardly shown has standard error
$\se \left( \widehat{\Delta_i^{(E \shp C)}} \right) = 
\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))-
q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))+
q_{ij}\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))+
q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\Delta_i^{[E \nat C]}$ as $\widehat{\Delta_i^{[E \nat C]}}=\widehat{\olep_i^{[E \nat C]}}-q_{ij}\widehat{\olep_j^{[E \nat C]}}=
\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]+
q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))],$ which one can straightforwardly show has standard error
$\sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$ 
and $s_2=\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]/\sqrt{M}$.
\item For $\Delta_i^{[EC]}$, we recall that $\olep_j^{[EC]}=0$, so $\Delta_i^{[EC]}=\olep_i^{[EC]}$. So we use
$\widehat{\olep_i^{[EC]}}$ for $\widehat{\Delta_i^{[EC]}}$.
\item Finally, we can also estimate $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ simply as $\widehat{\olr_{i \bs i}}$
since $\olr_{j \bs i}=0$.
\end{enumerate}
As for (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), it is again the case that (\ref{eq:decomp_overall_delta})
applies for the estimators as well as for the quantities being estimated.

\section{Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}

%***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.

In Fig. \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[EC]}$, and IGR $-\Delta_i^{[EC]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
From section \ref{sect:math_simplifications}, 
\begin{align}
\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
&= 0,
\end{align}
and
\begin{align}
\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
&= 0. %\label{eq:decomp:j:EnatC}
\end{align}
Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[EC]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}$ are zero so 
\begin{equation}
\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
\end{equation}
and 
\begin{equation}
\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}=\olep_i^{[EC]}
\end{equation}
Similarly for IGR $-\Delta_i^{[EC]}$, 
\begin{equation}
(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-(\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}) = \olr_{i \bs i}-\olep_i^{[EC]}.
\end{equation}
Thus $q_{ij}$ does not matter for anything computed for Fig. \ref{MT-fig:fig4}.

\clearpage
\newpage

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig3_qij.pdf}
\caption[Analogous to Fig. \ref{MT-fig:fig3} but making a different choice for $q_{ij}$]
{Analogous to Fig. \ref{MT-fig:fig3} of the main text, but using 
$q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. See the caption of
that figure for details. Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3qij,5)}, so are not displayed. \label{fig:fig3_alt}}
\end{figure}
%***DAN: Jasmin, in the caption of Fig. 3 in the main text there is a statment about standard
%error. We need another such statement here, but for the standard errors that applied here.

\clearpage
\newpage

%you store your bibliographic info in refs.bib, see that file
\bibliographystyle{ecology_letters2}
\bibliography{refs}

\end{document}