### 2023-07-03 paper/RefereeResponses1.tex
continued working on responses

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=1.5cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{hyperref}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\DeclareMathOperator{\E}{\mathbb{E}}% expected value
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\cor}{cor}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\prob}{P}

%attempt 1 at nat and sharp
%\newcommand{\nat}{\mathlarger{\natural}}
%\newcommand{\shp}{\mathlarger{\sharp}}

%attempt 2 at nat and sharp
%\newcommand{\nat}{\raisebox{1pt}{\mathsmaller{\mathsmaller{/\hspace{-2pt}/}}}}
%\newcommand{\shp}{\#}

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\bs}{\backslash}

%header material for paper
\title{Responses to referees for ``Asymmetric relationships and their effects on coexistence''}
\date{July 2023}

\begin{document}

\noindent \emph{Dear Editors,} \\

\noindent \emph{Thank you for the useful and very positive reviews of our manuscript, and for
the opportunity to resubmit a revised version. We have addressed all referee comments below and in
the manuscript, and hope the paper can now be rapidly accepted. } \\

\noindent \emph{We have replied to referee comments below in italic, with each response preceded
by ``***'', to make our responses easier to find.} \\

\noindent \emph{We submit the following pdf files: ...} \\

\noindent \emph{We additionally submit the following latex files, necessary for compiling the
main text latex: ...} \\

\noindent \emph{The document showing the changes we made does not include links such as citations and
references to figures, but it does accurately represent the changes made to the text. 
If all latex files are deposited into the same folder, it should then be possible to 
compile the main text latex from that folder.} \\

\noindent \emph{We again thank the referees and editors for their feedback and positive reviews!} \\

\noindent \emph{Yours,}

\noindent \emph{Dan Reuman and Jasmin Albert} \\

\noindent \textbf{Comments of referee 1:} \\

%The C) that precedes each comment is because we will later add numbers, C1, C2, etc., after making sure these
%comments are divided up to our liking. Then we can refer to each comment by its number, which can be useful.
\noindent C1) This is a very interesting paper that adds to our understanding of modern coexistence theory through the decomposition of the ways in which competitive pressure and environment can be correlated.\\

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C2) Specifically, the focus is on asymmetric tail association (ATA), by which two random variables with given, fixed marginal distributions and correlation can nevertheless differ in the way that their right and left tails are correlated. This is one among many ways that one could characterize or decompose the full bivariate distribution between competition and environment, but the authors argue and demonstrate that ATAs of this kind are (a) prevalent in natural systems and (b) can have a meaningful and sometimes qualitative effect on the outcome of coexistence between two species.  The combination of results for theoretical distirbutions but also for experimental data is very powerful and convincing. \\

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C3) I have one overarching question, which I think may be important for ensuring that those points (a) and (b) are robust and then some minor comments that may improve the readability of the manuscript for the EL audience. \\

\noindent ***\emph{response} \\

\noindent C4) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.

My first comment related to this is that some form of this text (whcih is relatively short---around half a page in the SI) and, importantly, the Figure should be in the main text.  I think this is such a key part of the definitions that are being used for the rest of the paper that readers will want to get an intuition for what the "partial sharp" distributions look like and how they are arrived at. \\

\noindent ***\emph{<Move SI Fig S1 and associated text back to the MT – at least two referees asked for it.>} \\

\noindent C5) Perhaps equally important to clarifying this process for the readership is whether this process is unique. I could not quite get a sense of whether there could be alternative versions of this removal of ATA from a bivariate distribution, that still satisfy the authors' constraints on marginal distributions and overall correlation between the two variables. Are there alternatives, or is the authors' procedure unique in achieving the desired outcome?  If the former, then I would want to understand if the outcomes downstream tend to be robust to the specific method used. If the latter, it would be helpful to have an explanation as to why this is a unique decomposition. \\

\noindent ***\emph{<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>} \\

\noindent C6) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
$E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.  Maybe this is standard notation, but somehow it may be clearer to put the subscript on the E for expectation value, to indicate that the probability distribution used to generate the definition of the averaging is what differs between the two expressions, while the variables E and C themselves are in the end just dummy variables to be integrated over.  \\

\noindent ***\emph{<This one asks for possibly changing notation. I don’t think we should do it, but we need to explain in a response what is actually going on, why we chose the notation that we do, and then add a new SM section addressing what the referee is talking about. Actually it’s the bivariate distribution that changes when you apply ||, tho the marginals are not changed. So the referee does have a bit of a point – E^{||} by itself is actually the same distribution as E, it’s just that it’s a different random variable in the sense of being related differently to C^{||} than E is to C. This is the same issue with the # notation that has already been established. It’s actually formally correct in the sense of random variables if we view *all* our random variables as being measurable functions from the *same* probability space. So in the end the referee does not really have a point if one wants to be formal. But you can explain all this in a new sup mat section.>} \\

\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).  These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species. Is this really a demonstration that ATAs occur for C and E?  If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).  Maybe I am missing something. \\

\noindent ***\emph{<Yes, the referee is right about this, that's what we're doing. Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it.>} \\

\noindent C8) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.  I think $r_{j \bs i}$ must also assume that $j$ is at steady state, but this isn't mentioned. Perhaps that could be added for completeness. \\

\noindent ***\emph{<Just do this and say we did it - it's a good enough idea.>} \\

\noindent C9) E.g. line 100 needs removal of the word "are" or the addition of the word "and" somewhere to make sense. I'd consider going through the paper carefully for any typos or small corrections of this kind. \\

\noindent ***\emph{response} \\

\noindent \textbf{Comments of referee 2:} \\

\noindent C10) In their paper the authors decompose the “storage effect” into two separate effects, what they call the “asymmetric tail association” and the covariance term per se. The authors explain well why they do this and why they think this may be important. Additionally, they show the best possible efforts to convince the readers why such a decomposition may be important and what we may learn from it. Overall, the paper is interesting and quite understandable. I think the authors have done a very good job. \\

\noindent ***\emph{Thank you for the very positive feedback!} \\

\noindent C11) Unfortunately, I’m personally not convinced that storage effect (and therefore probably ATA) are relevant for coexistence in nature and therefore I’m sceptical about this new approach. I want to be clear that this is, in part, my personal opinion and I know experts in the field which have a very different view. What the authors have shown (and it appears to be correct) is that in a model where niche differentiation via different resources and predators is impossible, then ATA can be an important driver of coexistence, and I agree. However, in every application where both storage effect, relative non-linearity and niche differentiation have been measured, there we see that storage effect and relative non-linearity where very small compared to simple resource differentiation. But, one might argue, that we have barely any such empirical applications, and I would agree, which is why I label this my personal opinion. (The few papers that I know of are \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1002%2Fecy.2726&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=HbUTjzCNQrvGSARY5u3Eqme88DbKiNyiAVNOUGzjcbE%3D&reserved=0}{Zepeda et al. 2019, Ecology}, \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1890%2F14-1741.1&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=KKIBuLjBN2vNGWfEpduAGGrMAt7ov9a4F53CmlZQlO0%3D&reserved=0}{Chu and Adler 2015}, and Ellner et al. 2019, see also the new paper by \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fesajournals.onlinelibrary.wiley.com%2Fdoi%2Fepdf%2F10.1002%2Fecm.1585&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=YaqsZ6axu6rYV12kh9IiRIbES7ZUMWIN7HXgB08iPzc%3D&reserved=0}{Stump and Vasseur}, which I have not yet read fully) In my understanding, ATA is a decomposition of storage effect into even more components, therefore I would expect, on average the ATA effect to be slightly smaller than storage effect (but, as you have shown can also be larger than storage effect). I would therefore assume that ATA is generally weak compared to niche differentiation (but again, admittedly based on few datapoints). Obviously, the authors may disagree with my personal view, in this case I’d like to see a discussion of this. \\

\noindent ***\emph{<Add a discussion, either in the Discussion or in an SM section called something like 
``Additional discussion points". Acknowledge the viewpoint, state that insufficient studies have been done to be sure
. Thank the referee for the openness that this is their own opinion - it's very welcome when a referee qualifies
what they say in this way which permits consideration of alternative viewpoints rather than just being forced to
do what referees say because they are in the position of power. Point out that we have not only worked with the lotery
model, as the referee mentions, we have also worked with an empirical example, and that one definitely supported
the importance of ATAs. Though we accept that future work may reveal that storage effects, and therefore probably
also ATAs, are not so important in field systems, we think it is also possible that this will not be 
revealed to be the case, and so we think it is important to not pre-judge the issue when (as the referee 
admits) there is insufficient data. Our results add another piece of evidence to the small collection of 
evidence as to the importance of storage effects, as well as adding the first evidence on the potential 
importance of ATAs. Also emphasize what the referee already seems to accept - that even if storage effects are 
small, ATA effects can be larger in magnitude if correltion effects per se counteract ATA effects.

Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms. For example, the temporal storage effect requires positive covariance between the effects of competition and the environment (Chesson and Huntly 1989): in years most favorable to a species, it should also be the most vulnerable to competition. However, we did not specify our vital rate regressions in a way that makes the covariation between the environment (or random year effects) and competition explicit. As a result, our model may not allow full expression of a temporal storage effect and the simulations may underestimate its strength. Such differences in model specification may explain why a previous analysis, based on a very different modeling approach, showed a strong storage effect in the Kansas community (Adler et al. 2006) in contrast to the weak effect we observed here. We plan to address this model specification issue in future work focused on the population and community consequences of environmental variability."

Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
to the Ellner approach of quantifying storage effects?

Finally, I notice that sometimes for our results, ATA effects and effects of correlation per se counter
each other. If that happens, storage effects could be weak, while ATA effects are still strong.

I suppose one must always accept the possibility that future work will render one's results less
meaningful or obsolete.
>} \\

\noindent C12) I like the conceptual idea of removing the long tail. \\

\noindent ***\emph{Thank you for the positive feedback.} \\

\noindent C13) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does. I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed. Additionally, it might be beneficial to indicate which distribution is then used to compute which part of the decomposition. For example, in Figure S1 d you show $E_i^{||}$, but from my understanding this distribution would be used to compute $E_i^{(EC)}$, correct? \\

\noindent ***\emph{<Clearly since two referees have now mentioned it we need to put some version of Fig S1 back into the
main text. The earlier referee wanted to see the whole figure brought back, but this referee wants to see just part, i.e.,
a simpler version, as specified. It's not clear to me how it would help the reader to just see the three panels the referee
suggests. Only 2 of the suggested panels are currently part of Fig. S1 - the suggestion of putting in the uncorrelated
distribution would be new. The current Fig. S1 serves the purpose of describing exactly the process of removing ATAs. 
The new suggested figure would not serve that purpose, it would instead maybe give the reader a sense for what these
distributions look like when manipulated in the ways we are talking about. But it's not real clear to me that much
intuition will be conveyed by doing that. One way to reply would be just to add a panel to the figure that shows 
the uncorrelated case, and move the resulting 5-panel figure back in. Then we'd explain that 1) the other referee asked
for the whole figure, 2) it's not clear to us that the simplified version serves any good purpose, it certainly 
does not serve the purpose of explaining the process of removing ATAs, but 3) we went ahead and added the uncorrelated
case anyway, since that's what was used previously, so it makes sense to have that point of comparison. If we do
this, Jasmin can add the new panel to the figure.>} \\

\noindent C14) You focus solely on storage effect with the argument that it’s the only where ATA can actually occur. But on a broader level you essentially decompose the EC distribution into it’s different momentums. One only carrying about the mean (i.e. first order correct), one with only carrying about the mean and covariance (i.e. getting the first two orders correct), and one with all other orders correct. Given this interpretation of your work, I might apply the same logic to the relative-nonlinearity, where I decompose the relative non-linearity into the variance term and the term with long tails. (one might also even decompose the actual distribution into more distributions, getting successively more and more momentums of the distribution correct, but I would argue that this is probably less useful). \\

\noindent ***\emph{We are not sure we understand precisely what the referee is saying here because their language
appears more suggestive than formally correct. Perhaps they mean
``moment'' rather than ``momentum''? Perhaps they mean ``caring'' instead of ``carrying''? I think the phrase 
``term with long tails'' does not precisely relate to what we're doing since marginal distributions
of the bivariate distribution $(E,C)$ are unchanged in all our approaches, and hence tails are unchanged. 
It's just relationships
between $E$ and $C$ that are modified in the different scenarios we consider. The referee's ideas, here,
if made precise, may be about potentially using a series of approximations to the bivariate distribution $(E,C)$
that have the same marginals as $(E,C)$ itself, but that get successively more co-moments right, compared to the
co-moments of $(E,C)$ itself (and the 
remaining co-moments would be zero, we presume, if
such distributions can be defined). Such an approach, applied to relative nonlinearity, would not give
anything interesting because relative nonlinearity only depends on the marginal distributions, which, 
under this framework, are always equal to the original marginals of $(E,C)$. On the other hand,
perhaps the referee means to take a series of approximations of $(E,C)$ so that the $n$th 
approximate distribution gets both moments and co-moments of order $\leq n$ right, with remaining
moments and co-moments zero (if such distributions can be defined). That may be interesting, but would
certainly be a different paper from ours, since our paper is about \emph{relationships} between $E$ and $C$
(while leaving the univariate distributions of $E$ and $C$ themselves fixed), whereas this idea of the 
referee would make approximations
both to the relationship between $E$ and $C$ and to the marginals, simultaneously. Nevertheless, 
we are pleased the referee 
is interested in our approach to the extent that they are generating potential ideas for future research.} \\

\noindent C15) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\

\noindent ***\emph{We modified the sentence.} \\

\noindent C16) L41: The “we” is slightly confusing because the author teams share only one author \\

\noindent ***\emph{We changed the wording here.} \\

\noindent C17) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\

\noindent ***\emph{There is no space to intro these things, so we limited this text to just mentioning 
extinction risks and stability of ecosystem functioning. Whereas some ecologists may not have heard of 
Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
a bell for most ecologists.} \\

\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\

\noindent ***\emph{We spent 15-20 minutes or so trying to add an explanation that was succinct enough to not break
the flow of the paper while still being informative, but we could not do it. Any more than a short sentence would break the 
flow, and we felt that all the one-sentence explanations we could come up with were unclear. 
FYI, in a nutshell the idea is as follows. Assume that low values of an environmental variable reduce the 
growth rates of local populations and high values elevate those growth rates. In that case one may
expect that left-tail associations increase metapopulation extinction risks because then environmental
``catastrophes'' are spatially synchronized, presumably reducing the potential for rescue effects. Whereas
``bonanzas'' (extremely good conditions, with a lot of growth) are much less correlated. The case where
high values of the variable reduce the growth and low values elevate it is similar, but it's right-tail association
that elevates extinction risk. This is most of the explanation, but it gets a bit more nuanced, still - 
see the paper for additional details. Thanks for your interest, by the way.} \\

\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\

\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
and main ideas of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
defined oihn Theory, which they are.} \\

\noindent C20) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters. \\

\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
paper possible, but, of the comments 
C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
and we have revised the text C19 mentions so as to robustly address that comment (see response above). 
The manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
the link to that other literature, so that readers can explore it if they so choose, 
and to show that ATAs have been found to be important in ecology, generally, to motivate the study.
Furthermore, we don't think our paper is any more specialized than the papers Ellner et al (2016, 2019) which
we cite, and which inspired our paper, and which were published in Ecology Letters. Nor is it more 
specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
central in ecology these days, and increasingly so. It makes no sense to us to sideline them into 
"specialist journals", which few people read, especially when they are about mainstream topics such
as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
paper broadly accessible.} \\
%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.

\noindent C21) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\

\noindent ***\emph{We replaced the reference.} \\
%Jasmin can you please do this?

\noindent C22) L74: The optimum might actually be the same, but they need different responses. \\

\noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\

\noindent C23) Eq. 1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\

\noindent ***\emph{We disagree, and would prefer not to make this change, for several reasons. 
First, the lottery model is a very simple model, at the same level of complexity
as models introduced in essentially all introductory 
undergraduate ecology courses (e.g., the Lotka-Volterra competition model), so there is no argument that 
the readership of Ecology Letters cannot easily understand the material presented here.
Second, this material serves a clear motivating purpose, most appropriately placed in the Introduction. 
The last author (DCR) has found, from decades of pitching 
quantitative papers to general ecological audiences, that clear motivation is essential - it is crucial to demonstrate 
clearly that an effect one plans to elaborate is present, and potentially important, prior to asking the reader to engage 
the kind of mental energy necessary to read the guts of the paper. So this 
paragraph is for showing that the effect of ATAs on coexistence can be real, at least for a simple 
example. That goal is convincingly 
accomplished by the paragraph. Given that the math is simple, and also more precise and definitive than 
qualitative verbal arguments that usually take up space in Introductions, we think this material is actually
more beneficial to the readership and more effective for the intended motivational purpose than would be more words
(and we see based on C25 below that the referee appears to agree with us here). 
Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
to want to exclude mathematics of any kind from the Introduction, especially when the math is simple,
accessible, and serves a clear motivational purpose, seems like a dated bias to us. 
Finally, we firmly believe that ecology as a discipline gets stronger as it continues the clear
trend of the last few decades of incorporating more quantitative approaches. We believe we
benefit the field by challenging assumptions and biases such as categorically not putting equations 
in the Introduction.
Overall, we think making this suggested change would make the paper weaker, and would pander to an
outmoded and unhelpful tradition. So we will not make the change unless the 
editor insists on it as a condition of publication.} \\

\noindent C24) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\

\noindent ***\emph{OK, we made this change, even though the same equation already appears just a few lines
above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\

\noindent C25) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\

\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
without first stating the lottery model, since the result uses that model. } \\

\noindent C) L126: Do you do this in the appendix? \\

\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
To avoid making claims that we have not fully supported
with rigorous mathematics or other formal methods, we rephrased to say the iid assumption is 
``probably not necessary''. Thanks for spurring us to do this.} \\

\noindent C) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\

\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
the $\#$ term because that terms involves both of them varying (though independently).} \\

\noindent C) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\

\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
to do this the first time we use the hard brackets. So thanks for pointing this out.
We do not know what the referee means by ``using curly under brackets'' so have not done anything
in response to that part of this comment.} \\

\noindent C) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work (and potential for error) 
of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
paper. And we could instead spend that time working on the concepts underlying the next paper.} \\

\noindent C) L316: So why do you show values of $\sigma >6$? And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA. In 1d it simply appears that the variance scales with the density. 1e is quite unconvincing to me. That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example? \\

\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
trying to do here is accurately portray our results about
when ATAs were important and when they were not. To us, the most forthright way of doing that is to
show \emph{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
our results seems important when such is our goal. Furthermore, there's no real cost to showing
all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
matters for continuing to explore when ATAs are important and when not is probably the strength
of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
based on such a thing.

Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
%the original paper and extract from p-value info, or we could come up with a better example,
%or something else, not sure, let's discuss.
} \\

\noindent C) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\

%Jasmin, isn't there already such a line? It just gets overplotted sometimes, right? But you can see it on
%panel (e), for instance, right? I've written a response assuming that, but let's briefly discuss.
\noindent ***\emph{There is already a line at $y=0$, you can see it, for instance, on panel (e). 
It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
for different $y$-axis extents is so that you can see the variation within each panel with maximum 
clarity. The main point, here, is to compare the relative contributions of different components
to coexistence, so it's important to show within-plot variation (between lines and across 
values of the $x$-axis) clearly. We thank the referee for expressing their own uncertainty, which was
useful to be aware of.} \\

\noindent C) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case. \\

\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
is usually one of the largest-magnitude terms. Our results as to the importance of ATAs for coexistence 
contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
of section 4.1.} \\

\noindent C) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\

\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
such errors. But we're happy to hear specific suggestions.} \\

\noindent \textbf{Comments of referee 3:} \\

\noindent C) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked. ATAs can give additional mechanisms that either promote or hinder coexistence in a similar fashion and magnitude to other previously recognised mechanisms such as the storage effect. To illustrate these ideas the authors combine results from mathematical models, simulations, and empirical examples, and show that under a certain combination of parameters ATAs can be a significant mechanism of species coexistence. \\

\noindent ***\emph{response} \\

\noindent C) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\

\noindent ***\emph{response} \\

\noindent C) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations. Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics. I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript. \\

\noindent ***\emph{response} \\

\noindent C) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences. \\

\noindent ***\emph{response} \\

\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, but how ATAs enter in play is not well solved. It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented. \\

\noindent ***\emph{response} \\

\noindent C) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor. Could the authors briefly indicate how this could occur? \\

\noindent ***\emph{response} \\

\noindent C) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\

\noindent ***\emph{response} \\

\noindent C) The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model. \\

\noindent ***\emph{response} \\

\noindent C) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system? this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones. \\

\noindent ***\emph{response} \\

\noindent \textbf{Comments of the editor:} \\

\noindent C) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\

\noindent ***\emph{response} \\

\noindent C) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists. My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing. The authors have to make an effort to use a more biological/ecological language with reference to real world potential examples/cases/systems. Some mathematical material could be moved to Supplementary material. \\

\noindent ***\emph{response} \\

\end{document}