### 2023-09-11 paper/RefereeResponses2.tex
started doc for second round of responses

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=1.5cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{hyperref}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\DeclareMathOperator{\E}{\mathbb{E}}% expected value
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\cor}{cor}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\prob}{P}

%attempt 1 at nat and sharp
%\newcommand{\nat}{\mathlarger{\natural}}
%\newcommand{\shp}{\mathlarger{\sharp}}

%attempt 2 at nat and sharp
%\newcommand{\nat}{\raisebox{1pt}{\mathsmaller{\mathsmaller{/\hspace{-2pt}/}}}}
%\newcommand{\shp}{\#}

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\bs}{\backslash}
\newcommand{\olep}{\overline{\epsilon}}

%header material for paper
\title{Responses to referees for ``Asymmetric relationships and their effects on coexistence''}
\date{September 2023}

\begin{document}

\noindent \emph{Dear Editors,} \\

\noindent \emph{Thank you for the useful and very positive reviews of our manuscript, and for
the opportunity to resubmit a revised version. We have addressed all referee comments below and in
the manuscript, and hope the paper can now be rapidly accepted. } \\

\noindent \emph{We have replied to referee comments below in italic, with each response preceded
by ``***'', to make our responses easier to find. We submit the following pdf files: The main text pdf, including all figures; 
the Supporting Information; and a version of the main text with changes tracked. The document showing 
the changes we made does not include links such as citations and
references to figures, but it does accurately represent the changes made to the text.} \\

\noindent \emph{We again thank the referees and editors for their feedback and positive reviews!} \\

\noindent \emph{Yours,}

\noindent \emph{Dan Reuman and Jasmin Albert} \\
%DONE with cover letter

\noindent \textbf{Comments of referee 1:} \\

\noindent C1) First, to consider introducing a discussion of the robustness to more esoteric definitions of the partial-sharp distributions, and adding this briefly to the main text.  Primarily as the authors already suggest in their response letter, as a signpost for future work.\\

\noindent ***\emph{} \\
%DONE

\noindent C2) The only other minor suggestion is that the authors clarify that in Fig 1 they are simply showing that ATAs do occur in nature.  I was not suggesting that the authors include examples pertinent to their question in Figure 1---as they say, readers may not be expecting this. I don't think that the connection between the ATAs in Fig 1 and the rest of the project are that clear, as I already expressed. But I respect the authors wishes (and their other feedback) that it is worthwhile including these. Just a sentence or so at some point explaining that these are not directly related to the ATAs relevant in the remainder of the paper may avoid confusion. \\

\noindent ***\emph{} \\

\noindent \textbf{Comments of referee 2:} \\

\noindent C3) This is a novel contribution highlighting that asymmetric tail associations in correlations between two variables was a very simple aspect not properly considered before but with huge implications for the maintenance of biodiversity. I congratulate the authors for their work. \\

\noindent ***\emph{Thank you for the very positive feedback!} \\
%DONE

\noindent \textbf{Comments of referee 3:} \\

\noindent C4) I thank the authors for the change and the nice discussion/replies to my questions. I especially like the new figure 2 which greatly helped my understanding of how to remove ATA. \\

\noindent ***\emph{Many thanks for the positive feedback!} \\

\noindent \textbf{Comments of the editor:} \\

\noindent C5) Reviewers are enthusiastic about this paper demonstrating the importance of how variable relationships depending on when they are weak or strong influence competitive exclusion and coexistence. \\

\noindent ***\emph{Thank you for pointing out.} \\

\noindent C6) The legends of figures are rather long and dense. Figures should be self-explanatory. I suggest adding to all of them the sentence at the end: “See Table 1 for used annotations and abbreviations.” \\

\noindent ***\emph{We }

\noindent C7) The legend in Fig. 1 is too long and has variables not present in the graphs but mentioned in other sections of the paper. For example, in line 44-45: “The variables (B1, B2) described in the Introduction…”. I suggest revising if this information is needed at all in this Figure. \\

\noindent ***\emph{Thank you for the very positive feedback!} \\

\noindent C8) Main concepts should be described in the introduction in a way that could be understood for a broad audience. In this sense, I suggest adding a short non- mathematical definition for storage effects even if you go in more detail in the Theory section in lines 59-60 \\

\noindent ***\emph{Thank you for the very positive feedback!} \\

\noindent C9) Page 2, Lines 59-60 “coexistence into the contributions of various mechanisms, with names such as storage effects (Chesson1994). Storage effects refer to….. Storage effects are reviewed conceptually below and defined formally in Theory. \\

\noindent ***\emph{Thank you for the very positive feedback!} \\

\noindent C10)  Page 2, Line 60: Start a new paragraph with the sentence “MCT has been applied to several systems…” \\

\noindent ***\emph{Thank you for the very positive feedback!} \\

\emph{We thank the referees and editor for spurring us to make these changes, as the paper is improved in
its accessibility. The paper is certainly now no less accessible than many other papers I've read recently in 
Ecology Letters, including the two Ellner et al. papers (cited in our study) which helped inspire this 
work and which were published in Ecology Letters.} 
%DONE

\end{document}